The necessity of hypermedia RDF and an approach to achieve it

### 2012-03-01 eswc-2012/workshop2.tex
hypermedia RDF
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@138 43ef1190-90fd-4dd7-a769-d0121f02b80e

Mike Amundsen defines hypermedia types\footnote{\url{http://amundsen.com/hypermedia/}} as 
\begin{quote}
Hypermedia Types are MIME media types that contain native
hyper-linking semantics that induce application flow. For example,
HTML is a hypermedia type; XML is not.
\end{quote}
Furthermore, he defines a classification scheme called H~Factor as ``a
measurement of the level of hypermedia support and sophistication of a
media-type.'' The REST & WOA Wiki defines ``the Hypermedia
Scale''\footnote{\url{http://restpatterns.org/Articles/The_Hypermedia_Scale}},
where the categorization is based on capabilities for create, read,
update and delete operations.  Standing on it's own, RDF is a
hypermedia type, but only at the \textsf{LO} and \textsf{CL} levels,
and just an R~Type (read) on the Hypermedia Scale. We aim at improving
this situation.

@@

>>>>

Mike Amundsen defines hypermedia types\footnote{\url{http://amundsen.com/hypermedia/}} as 
\begin{quote}
Hypermedia Types are MIME media types that contain native
hyper-linking semantics that induce application flow. For example,
HTML is a hypermedia type; XML is not.
\end{quote}
Furthermore, he defines a classification scheme called H~Factor as ``a
measurement of the level of hypermedia support and sophistication of a
media-type.'' The REST \& WOA Wiki defines ``the Hypermedia
Scale''\footnote{\url{http://restpatterns.org/Articles/The\_Hypermedia\_Scale}},
where the categorization is based on capabilities for create, read,
update and delete operations.  Standing on it's own, RDF is a
hypermedia type, but only at the \textsf{LO} and \textsf{CL} levels,
and just an R~Type (read) on the Hypermedia Scale. We aim at improving
this situation.

@@

\section{Defining hypermedia RDF}

We noted that for a read-write RESTful protocol, we need to say in the
RDF message itself what can of operations can be made. We also note
that only information resources can be manipulated, and we argued that
it should be possible to add the required triples to every
resource. We should be able to define hypermedia RDF in terms of a
minimal vocabulary, but like the H~Factor web page, we shall find it
instructive to use examples. In the following, we use Turtle syntax
with prefixes omitted for brevity. Also, we note that in many cases,
we are making statements about the current resource, which given a
reasonable base URI can be written as \texttt{<>}. These are the H~Factors:

\begin{description}
\item[LE] Support for embedded links (HTTP GET)

Not relevant for RDF.

\item[LO] Support for out-bound navigational links (HTTP GET)

Trivially supported by RDF.

\item[LT] Support for templated queries (HTTP GET)

Undefined by this paper. Possibly supported by RDForms\footnote{\url{http://vocab.deri.ie/rdforms}}?

\item[LN] Support for non-idempotent updates (HTTP POST)

\begin{verbatim}
<> hm:canBe hm:mergedInto ;
   hm:createSimilarAt <../> .
\end{verbatim}

The first triple says that the current resource can be merged with
another resource. In a typical HTTP case, this would be achieved by
POSTing to the current resource with an RDF payload, and the server
would perform a RDF merge of the payload with the current resource.

The second triple exists to make it possible (in the typical HTTP
case) to POST an RDF payload to some URI and the server will itself
assign a URI to the posted RDF payload.

\item[LI] Support for idempotent updates (HTTP PUT, DELETE)

\begin{verbatim}
<> hm:canBe hm:replaced ,
            hm:deleted .
\end{verbatim}

The first triple says that the current resource may be replaced by, in
the HTTP case, PUTting an RDF payload to the resource URI. The second
triple says that the resource may be deleted, typically by a HTTP
DELETE on the resource URI.

\item[CR] Support for modifying control data for read requests
  (e.g. HTTP Accept-* headers).

\begin{verbatim}
<> hm:acceptsFormats <http://www.w3.org/ns/formats/Turtle> .
\end{verbatim}

\item[CU] Support for modifying control data for update requests
  (e. g. Content-* headers).

Undefined by this paper.

\item[CM] Support for indiciating the interface method for requests
  (e.g. HTTP GET, POST, PUT, DELETE methods).

E.g.

\begin{verbatim}
hm:replaced        hm:httpMethod "PUT" .
hm:deleted         hm:httpMethod "DELETE" .
hm:mergedInto      hm:httpMethod "POST" .
hm:createSimilarAt hm:httpMethod "POST" .
\end{verbatim}

\item[CL] Support for adding semantic meaning to link elements using
  link relations (e.g. HTML rel attribute).

Trivially supported by RDF.

\end{description}

We see that with a simple vocabulary, RDF can become a very powerful
hypermedia type. We also note that this vocabulary does enable agents
to do the same operations as the SPARQL  1.1 Graph Store HTTP
Protocol\footnote{\url{http://www.w3.org/TR/2011/WD-sparql11-http-rdf-update-20110512/}},
but being fully RESTful. The specification is currently not RESTful
per the discussion above, even though it was a key design goal. It
should also be possible to use this on the default (nameless) graph by
assigning it a name in the service description rather than defining it
in a out-of-band specification like it is currently being done.

\bibliographystyle{plain-csmin}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
\bibliography{federation,webarch,webframeworks}

\end{document}
